Introduction to Structural Causal Models

### 2021-05-21 presentation_video.Rmd
n

# Foundations of SCMs

- Flexible simulations for higher order problems (intervention, counterfactual) based on elementary noise variables
- Graphical visualization via directed acyclic graph
- Nonparametric SEM 
- System of equations with functions
- **Assignments** ':=' (non-symmetric) 
- algebraic equation '=' (symmetry)
- **Error in Regression:** Omittable outside factor 
- **Error in SCM:** Latent (influential) factor that is pivotal for the model

@@

# Assumptions (1)

Independence:

@@

# Fundamental Differences (1)

@@

- First tow hold for Counterfactuals in SCM
- third not testable but can check via backdoor criterion in SCM [@hardtrecht]

@@

# Does and Do nots

@@

* ...only have conditional probabilities 
* ...are only interested in association 
* ...have very limited knowledge of your actual underlying causal structure

@@

* ... have a good understanding of underlying latent factors 
* ... have a lot of information on your underlying relationship in your data 
* ... are dealing with high stake settings 
* ... want to simulate different interventions before implementing them

@@

>>>>

* Topic 1: **Structural Causal Models**
* Topic 2: Linking d-separation with (conditional) independence & Causal Structure Learning
* Topic 3: do-calculus: estimating the effect of interventions & counterfactuals 
* Topic 4: Rubin’s Potential Outcome Framework & Causal Trees
* Topic 5: Algorithmic Recourse 
* Topic 6: Causal Interpretation of Black-Box Models

# Foundations of SCMs (1)

- Flexible simulations for higher order problems (intervention, counterfactual) 
- based on (elementary) 'noise'/latent variables
- Graphical visualization via directed acyclic graph
- Nonparametric SEM 
- System of equations with functions

# Foundations of SCMs (2)

- **Assignments** ':=' (non-symmetric) 
- algebraic equation '=' (symmetry)
- **Error in Regression:** Omittable outside factor 
- **Error in SCM:** Latent (influential) factor that is pivotal for the model

@@

# Assumptions - Independence

@@

# Fundamental Differences

@@

- First tow hold for Counterfactuals in SCM
- third not testable but can check via backdoor criterion in SCM
- Source: [@hardtrecht]

@@

# Dos and Do nots

@@

* ...only have conditional probabilities 
* ...are only interested in association 
* ...have very limited knowledge of your actual underlying structure

@@

* ... have an understanding of underlying relationships in the data
* ... are dealing with high stake settings 
* ... want to simulate different interventions

@@

# Questions:

Thank you so much for your time!

If you have any questions contact me via:

* Mattermost: daniel.saggau
* Email: daniel.saggau@campus.lmu.de
